### 2021-08-27 tlf-planning.Rmd

# Table for Planning

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize important deviations in the section 10.2, protocol deviations of a CSR.

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy analysis 
in the section 11.1, data sets analysed of a CSR.

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize critical demographic and baseline characteristics of the patients
in the section 11.2, demographic and other baseline characteristics of a CSR.

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize measurements of compliance of the patients
in the section 11.3, measurements of treatment compliance of a CSR.

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize primary and secondary efficacy endpoints. 
in the section 11.4, efficacy results and tabulations of individual patient data of a CSR.