### 2021-08-28 tlf-population.Rmd
add analysis population example

# Table for Study Patients {#population}

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy analysis 
in the section 11.1, data sets analysed of a CSR.

```{r}
library(haven) # Read SAS data
library(dplyr) # Manipulate data
library(tidyr) # Manipulate data
library(r2rtf) # Reporting in RTF format
```

In this chapter, we illustrate how to create a summary table for
analysis population in a study.

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_pop.pdf")
```

The first step is to read relevant datasets into R. 
For analysis population table, 
all the required information is saved in the ADSL dataset.
We can use `haven` package to read the dataset.

```{r}
adsl <- read_sas("adam_data/adsl.sas7bdat")
```

We illustrate how to prepare a report data for a simplified 
analysis population table using variables below:

- USUBJID: Unique subject identifier 
- ITTFL: Intent-to-treat population flag
- EFFFL: Efficacy population flag
- SAFFL: Safty population flag

```{r}
adsl %>%
  select(USUBJID, ITTFL, EFFFL, SAFFL) %>%
  head(4)
```
## Helper Functions

Before we write analysis code, let's discuss the possibility to
reuse R code by writing some toy helper function.

As discussed in 
[R for data science](https://r4ds.had.co.nz/functions.html#when-should-you-write-a-function),
"You should consider writing a function 
whenever you’ve copied and pasted a block of code more than twice"

In Chapter \@ref(dispositon), there are few repeating steps:

- format percentage using `formatC` 
- calculate number and percentage by groups

We create two ad-hoc functions and apply them to all rest of tables.

To format number and percentage, we created a function called `fmt_num`.
It is a very simple function that is a wrapper of `formatC`.

```{r}
fmt_num <- function(x, digits, width = digits + 4){
  formatC(x,
          digits = digits, 
          format = "f", 
          width = width
    )
}
```

The main reason to create `fmt_num` is to enhance readability of the analysis code.

For example, we can compare the two versions of code to format percentage 
used in \@ref(dispositon).

```{r, eval = FALSE}
formatC(n / n() * 100,
      digits = 1, format = "f", width = 5
    )
```

```{r, eval = FALSE}
fmt_pct(n / n() * 100)
```

To calculate number and percentage of patients by groups, 
we provide a simple (but not robust) wrapper function, `count_by` using `dplyr` and `tidyr`.

The function can be enhanced in multiple ways, but here we only focus on simplificy and readability.
More details about writting R functions can be found in [Stat454 course](https://stat545.com/functions-part1.html).

```{r}
count_by <- function(data,               # Input data set
                     grp,                # Group variable
                     var,                # Analysis variable
                     var_label = var,    # Analysis variable label
                     id = "USUBJID") {   # Subject ID variable 
  data <- data %>% rename(grp = !!grp, var = !!var, id = !!id)

  left_join(
    count(data, grp, var),
    count(data, grp, name = "tot"), 
    by = "grp",
  ) %>%
    mutate(
      pct = fmt_num(100 * n / tot, digits = 1),
      n   = fmt_num(n, digits = 0), 
      npct = paste0(n , " (", pct, ")")
    ) %>%
    pivot_wider(
      id_cols = var,
      names_from = grp,
      values_from = c(n, pct, npct),
      values_fill = list(n = "0", pct = fmt_pct(0))
    ) %>% 
    mutate(var_label = var_label)
}
```

By using `count_by`, we can simplify the analysis code as below.

```{r}
count_by(adsl, "TRT01PN", "EFFFL")
```

### Analysis Code

With the helper functions `count_by`, we can easily prepare report dataset as

```{r}
# Derive a randomization flag 
adsl <- adsl %>% mutate(RANDFL = "Y") 
pop <- count_by(adsl, "TRT01PN", "RANDFL", var_label = "Participants in Population") %>% 
          select(var_label,starts_with("n_"))
```

```{r}
pop1 <- bind_rows(
  count_by(adsl, "TRT01PN", "ITTFL", 
           var_label = "Participants included in ITT population"), 
  count_by(adsl, "TRT01PN", "EFFFL", 
           var_label = "Participants included in efficacy population"),
  count_by(adsl, "TRT01PN", "SAFFL", 
           var_label = "Participants included in safety population")
) %>% 
  filter(var == "Y") %>% 
  select(var_label, starts_with("npct_"))
```

```{r}
names(pop) <- gsub("n_", "npct_", names(pop))
tbl_pop <- bind_rows(pop, pop1)
tbl_pop
```

```{r}
rel_width <- c(2, rep(1, 3))
colheader <- " | \\line Placebo | Xanomeline \\line Low Dose| Xanomeline \\line High Dose"
tbl_pop %>%

  # Table title
  rtf_title("Participants Accounting in Analysis Population", 
            "(All Participants Randomized)") %>%

  # First row of column header
  rtf_colheader(colheader,
    col_rel_width = rel_width
  ) %>%

  # Second row of column header
  rtf_colheader(" | n (%) | n (%) | n (%)",
    border_top = "",
    col_rel_width = rel_width) %>%

  # Table body
  rtf_body(
    col_rel_width = rel_width,
    text_justification = c("l", rep("c", 3))
  ) %>%

  # Encoding RTF syntax
  rtf_encode() %>%

  # Save to a file
  write_rtf("tlf/tbl_pop.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_pop.pdf")
```