### 2021-08-28 tlf-population.Rmd
tlf-population: fix typo

# Table for Analysis Population {#population}

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy analysis 
in the section 11.1, data sets analysed of a CSR.

```{r}
library(haven) # Read SAS data
library(dplyr) # Manipulate data
library(tidyr) # Manipulate data
library(r2rtf) # Reporting in RTF format
```

In this chapter, we illustrate how to create a summary table for
analysis population in a study.

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_pop.pdf")
```

The first step is to read relevant datasets into R. 
For analysis population table, 
all the required information is saved in the ADSL dataset.
We can use `haven` package to read the dataset.

```{r}
adsl <- read_sas("adam_data/adsl.sas7bdat")
```

We illustrate how to prepare a report data for a simplified 
analysis population table using variables below:

- USUBJID: Unique subject identifier 
- ITTFL: Intent-to-treat population flag
- EFFFL: Efficacy population flag
- SAFFL: Safty population flag

```{r}
adsl %>%
  select(USUBJID, ITTFL, EFFFL, SAFFL) %>%
  head(4)
```
## Helper Functions

Before we write analysis code, let's discuss the possibility to
reuse R code by writing some toy helper function.

As discussed in 
[R for data science](https://r4ds.had.co.nz/functions.html#when-should-you-write-a-function),
"You should consider writing a function 
whenever you’ve copied and pasted a block of code more than twice"

In Chapter \@ref(disposition), there are few repeating steps to:

- format percentage using `formatC` 
- calculate number and percentage by groups

We create two ad-hoc functions and apply them to all rest of tables.

To format number and percentage, we created a function called `fmt_num`.
It is a very simple function that is a wrapper of `formatC`.

```{r}
fmt_num <- function(x, digits, width = digits + 4) {
  formatC(x,
    digits = digits,
    format = "f",
    width = width
  )
}
```

The main reason to create `fmt_num` is to enhance readability of the analysis code.

For example, we can compare the two versions of code to format percentage 
used in \@ref(disposition) and `fmt_num`.

```{r, eval = FALSE}
formatC(n / n() * 100,
  digits = 1, format = "f", width = 5
)
```

```{r, eval = FALSE}
fmt_num(n / n() * 100, digits = 1)
```

To calculate number and percentage of patients by groups, 
we provide a simple (but not robust) wrapper function, `count_by` using `dplyr` and `tidyr`.

The function can be enhanced in multiple ways, but here we only focus on simplificy and readability.
More details about writting R functions can be found in [Stat454 course](https://stat545.com/functions-part1.html).

```{r}
count_by <- function(data, # Input data set
                     grp, # Group variable
                     var, # Analysis variable
                     var_label = var, # Analysis variable label
                     id = "USUBJID") { # Subject ID variable
  data <- data %>% rename(grp = !!grp, var = !!var, id = !!id)

  left_join(
    count(data, grp, var),
    count(data, grp, name = "tot"),
    by = "grp",
  ) %>%
    mutate(
      pct = fmt_num(100 * n / tot, digits = 1),
      n = fmt_num(n, digits = 0),
      npct = paste0(n, " (", pct, ")")
    ) %>%
    pivot_wider(
      id_cols = var,
      names_from = grp,
      values_from = c(n, pct, npct),
      values_fill = list(n = "0", pct = fmt_pct(0))
    ) %>%
    mutate(var_label = var_label)
}
```

By using `count_by`, we can simplify the analysis code as below.

```{r}
count_by(adsl, "TRT01PN", "EFFFL") %>% 
  select(- ends_with(c("_54", "_81")))
```

## Analysis Code

With the helper functions `count_by`, we can easily prepare report dataset as

```{r}
# Derive a randomization flag
adsl <- adsl %>% mutate(RANDFL = "Y")

pop <- count_by(adsl, "TRT01PN", "RANDFL", 
                var_label = "Participants in Population") %>%
  select(var_label, starts_with("n_"))
```

```{r}
pop1 <- bind_rows(
  count_by(adsl, "TRT01PN", "ITTFL",
    var_label = "Participants included in ITT population"
  ),
  count_by(adsl, "TRT01PN", "EFFFL",
    var_label = "Participants included in efficacy population"
  ),
  count_by(adsl, "TRT01PN", "SAFFL",
    var_label = "Participants included in safety population"
  )
) %>%
  filter(var == "Y") %>%
  select(var_label, starts_with("npct_"))
```

Now we combined individual rows into the whole table for reporting purpose. 
`tbl_pop` is used as input for `r2rtf` to create final report.

```{r}
names(pop) <- gsub("n_", "npct_", names(pop))
tbl_pop <- bind_rows(pop, pop1)

tbl_pop %>% select(var_label, npct_0)
```

We start to define the format of the output.

```{r}
rel_width <- c(2, rep(1, 3))
colheader <- " | Placebo | Xanomeline line Low Dose| Xanomeline line High Dose"
tbl_pop %>%
  # Table title
  rtf_title(
    "Participants Accounting in Analysis Population",
    "(All Participants Randomized)"
  ) %>%
  # First row of column header
  rtf_colheader(colheader,
    col_rel_width = rel_width
  ) %>%
  # Second row of column header
  rtf_colheader(" | n (%) | n (%) | n (%)",
    border_top = "",
    col_rel_width = rel_width
  ) %>%
  # Table body
  rtf_body(
    col_rel_width = rel_width,
    text_justification = c("l", rep("c", 3))
  ) %>%
  # Encoding RTF syntax
  rtf_encode() %>%
  # Save to a file
  write_rtf("tlf/tbl_pop.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_pop.pdf")
```