### 2021-08-29 tlf-ae-summary.Rmd
place holder for ae tables

# AE Summary

Following [ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy analysis 
in the section 11.1, data sets analysed of a CSR.

```{r}
library(haven) # Read SAS data
library(dplyr) # Manipulate data
library(tidyr) # Manipulate data
library(r2rtf) # Reporting in RTF format
```