### 2021-09-13 intro.Rmd
Changes on Intro

# Introduction {#intro}

## Folder structure

In clinical trial development, developers needs to organize source code to deliver Study Data Tabulation Model (SDTM) or Analysis Dataset Model (ADaM) datasets and tables, listings and figures (TLFs).
For example, a Phase 3 clinical trial typically require hundreds of TLFs. 
A consistent and well-defined folder structure is crucial to manage a clinical trial analysis and reporting (A&R) project.

We recommend to use R package folder structure to organize all the A&R related source code and documentation for a clinical trial A&R project. 
By using the R package folder structure, you can follow the convention to organize your code as other contributors on CRAN.
This consistent approach simplifies the communication for all developers within and across organizations.

- For a new R developer, it is an essential step to learn R package when you want to share your work with others. 
You will learn one folder structure that is widely used in R community with outstanding tutorial and tools for free.

- For an experienced R developer, there is minimal learning curve.

- For an organization, it simplifies process, tool, template, and training development, because a unified folder structure is used to develop and maintain standard tool and analysis project.

The workflow around an R package can also improve the tractability and reproducibility for an analysis project [@marwick2018packaging].

In addition, R package folder structure are also recommended to develop Shiny app as discussed in [Chapter 20 of Master Shiny](https://mastering-shiny.org/scaling-packaging.html) book and [Engineering Production-Grade Shiny Apps](https://engineering-shiny.org/golem.html) book.

## In this book

This book is an intermediate level book by assuming reader has equipped some R programming and clinical development knowledge.
The assumption of each part is as below:

- Part 1, "Delivering TLFs in CSR" provides general information with examples to create tables, listings and figures. 
In this part, we assume readers are individual contributors of a clinical project with some experience in R programming. 
We expect readers are familiar with data manipulation in R. Some good reference includes [Hands-On Programming with R](https://rstudio-education.github.io/hopr/), 
[R for Data Science](https://r4ds.had.co.nz/) and [Data Manipulation with R](https://www.amazon.com/Data-Manipulation-R-Use/dp/0387747303).

- Part 2, "Clinical Trial Project" provides general information with examples to manage a clinical trial A&R 
project. In this part, we assume a reader is a project lead with experience in R package development. 
Some good reference includes [R packages](https://r-pkgs.org/) and [The tidyverse style guide](https://style.tidyverse.org/).

- Part 3, "eCTD Submission package" (maybe provide the full spelling for "eCTD" here) provides general information in preparing 
submission package related to clinical study report (CSR) in 
electronic Common Technical Document (eCTD). 
In this part, we assume a reader is a project lead of a clinical project 
with experience in R package development as well.

## Philosophy

We share the same philosophy described in [Section 1.1 of R Packages book](https://r-pkgs.org/intro.html?q=Marwick#intro-phil) and quote here.

- "Anything that can be automated, should be automated." 
- "Do as little as possible by hand. Do as much as possible with functions." 
- "The goal is to spend your time thinking about what you want to do 
rather than thinking about the minutiae of package structure."