### 2021-09-15 tlf-population.Rmd
update tlf-population

---
editor_options: 
  markdown: 
    wrap: 72
---

# Analysis Population {#population}

Following [ICH E3
guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize which patients were included in each efficacy
analysis in Section 11.1, Data Sets Analysed.

```{r}
library(haven) # Read SAS data
library(dplyr) # Manipulate data
library(tidyr) # Manipulate data
library(r2rtf) # Reporting in RTF format
```

In this chapter, we illustrate how to create a summary table for
analysis population in a study.

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_pop.pdf")
```

The first step is to read relevant datasets into R. For the analysis
population table, all the required information is saved in the ADSL
dataset. We can use the `haven` package to read the dataset.

```{r}
adsl <- read_sas("adam_data/adsl.sas7bdat")
```

We illustrate how to prepare a report data for a simplified analysis
population table using variables below:

-   USUBJID: Unique subject identifier
-   ITTFL: Intent-to-treat population flag
-   EFFFL: Efficacy population flag
-   SAFFL: Safty population flag

```{r}
adsl %>%
  select(USUBJID, ITTFL, EFFFL, SAFFL) %>%
  head(4)
```

## Helper Functions

Before we write analysis code, let's discuss the possibility to reuse R
code by writing some toy helper functions.

As discussed in [R for data
science](https://r4ds.had.co.nz/functions.html#when-should-you-write-a-function),
"You should consider writing a function whenever you've copied and
pasted a block of code more than twice".

In Chapter \@ref(disposition), there are few repeating steps to:

-   Format the percentages using the `formatC()` function.
-   Calculate the numbers and percentages by groups.

We create two ad-hoc functions and apply them to all the rest of the
tables.

To format numbers and percentages, we created a function called
`fmt_num()`. It is a very simple function wrapping `formatC()`.

```{r}
fmt_num <- function(x, digits, width = digits + 4) {
  formatC(x,
    digits = digits,
    format = "f",
    width = width
  )
}
```

The main reason to create the `fmt_num()` function is to enhance the
readability of the analysis code.

For example, we can compare the two versions of code to format the
percentage used in \@ref(disposition) and `fmt_num`.

```{r, eval = FALSE}
formatC(n / n() * 100,
  digits = 1, format = "f", width = 5
)
```

```{r, eval = FALSE}
fmt_num(n / n() * 100, digits = 1)
```

To calculate the numbers and percentages of patients by groups, we
provide a simple (but not robust) wrapper function, `count_by()` using
the `dplyr` and `tidyr` package.

The function can be enhanced in multiple ways, but here we only focus on
simplicity and readability. More details about writing R functions can
be found in the [Stat454
course](https://stat545.com/functions-part1.html).

```{r}
count_by <- function(data, # Input data set
                     grp,  # Group variable
                     var,  # Analysis variable
                     var_label = var, # Analysis variable label
                     id = "USUBJID") { # Subject ID variable
  data <- data %>% rename(grp = !!grp, var = !!var, id = !!id)

  left_join(
    count(data, grp, var),
    count(data, grp, name = "tot"),
    by = "grp",
  ) %>%
    mutate(
      pct = fmt_num(100 * n / tot, digits = 1),
      n = fmt_num(n, digits = 0),
      npct = paste0(n, " (", pct, ")")
    ) %>%
    pivot_wider(
      id_cols = var,
      names_from = grp,
      values_from = c(n, pct, npct),
      values_fill = list(n = "0", pct = fmt_num(0, digits = 0))
    ) %>%
    mutate(var_label = var_label)
}
```

By using the `count_by()` function, we can simplify the analysis code as
below.

```{r}
count_by(adsl, "TRT01PN", "EFFFL") %>% 
  select(- ends_with(c("_54", "_81")))
```

## Analysis Code

With the helper functions `count_by`, we can easily prepare a report
dataset as

```{r}
# Derive a randomization flag
adsl <- adsl %>% mutate(RANDFL = "Y")

pop <- count_by(adsl, "TRT01PN", "RANDFL", 
                var_label = "Participants in Population") %>%
  select(var_label, starts_with("n_"))
```

```{r}
pop1 <- bind_rows(
  count_by(adsl, "TRT01PN", "ITTFL",
    var_label = "Participants included in ITT population"
  ),
  count_by(adsl, "TRT01PN", "EFFFL",
    var_label = "Participants included in efficacy population"
  ),
  count_by(adsl, "TRT01PN", "SAFFL",
    var_label = "Participants included in safety population"
  )
) %>%
  filter(var == "Y") %>%
  select(var_label, starts_with("npct_"))
```

Now we combine individual rows into the whole table for reporting
purposes. `tbl_pop` is used as input for `r2rtf` to create the final
report.

```{r}
names(pop) <- gsub("n_", "npct_", names(pop))
tbl_pop <- bind_rows(pop, pop1)

tbl_pop %>% select(var_label, npct_0)
```

We start to define the format of the output.

```{r}
rel_width <- c(2, rep(1, 3))
colheader <- " | Placebo | Xanomeline line Low Dose| Xanomeline line High Dose"
tbl_pop %>%
  # Table title
  rtf_title(
    "Participants Accounting in Analysis Population",
    "(All Participants Randomized)"
  ) %>%
  # First row of column header
  rtf_colheader(colheader,
    col_rel_width = rel_width
  ) %>%
  # Second row of column header
  rtf_colheader(" | n (%) | n (%) | n (%)",
    border_top = "",
    col_rel_width = rel_width
  ) %>%
  # Table body
  rtf_body(
    col_rel_width = rel_width,
    text_justification = c("l", rep("c", 3))
  ) %>%
  # Encoding RTF syntax
  rtf_encode() %>%
  # Save to a file
  write_rtf("tlf/tbl_pop.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tbl_pop.pdf")
```

The procedure to generate an analysis population table can be summarized
as follow:

-   Step 1: Read data into R, i.e., `adsl`.
-   Step 2: Bind the counts/percentages of the ITT population, the
    efficacy population and the safety population by row using the
    `count_by()` function.
-   Step 3: Format the output in Step 2 by `r2rtf`.