### 2021-09-15 tlf-baseline.Rmd
update baseline table

# Baseline Characteristics

Following [ICH E3
guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf),
we need to summarize critical demographic and baseline characteristics
of the patients in Section 11.2, Demographic and Other Baseline
Characteristics.

In this chapter, we illustrate how to create a simplified baseline
characteristics table in a study.

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_base.pdf")
```

There are many R packages that can efficiently summarize baseline
information. The [`table1`](https://github.com/benjaminrich/table1) R
package is one of them.

```{r}
library(table1)
library(r2rtf)
library(haven)
library(dplyr)
library(tidyr)
library(stringr)
library(tools)
```

As in previous chapters, we first read `adsl` dataset that contain all
required information for baseline characteristics table.

```{r}
adsl <- read_sas("adam_data/adsl.sas7bdat")
```

For simplicity, we only analyze `SEX`, `AGE` and, `RACE` in this example
using the `table1` R package. More details of the `table` R package can
be found in the package
[vignettes](https://benjaminrich.github.io/table1/vignettes/table1-examples.html).

The `table1` R package directly creates an HTML report.

```{r}
ana <- adsl %>% 
  mutate(SEX = factor(SEX, c("F", "M"), c("Female", "Male")), 
         RACE = toTitleCase(tolower(RACE)))

tbl <- table1(~ SEX + AGE + RACE | TRT01P, data = ana)
tbl
```

The code below transfer the output into a data frame that only contain
ASCII character recommended by regulatory agencies. `tbl_base` is used
as input for `r2rtf` to create a final report.

```{r}
tbl_base <- tbl %>% 
              as.data.frame() %>% 
              as_tibble() %>%
              mutate( across(everything(), 
                      ~ str_replace_all(.x, intToUtf8(160), " ") ) 
                    )

names(tbl_base) <- str_replace_all(names(tbl_base), intToUtf8(160), " ")
tbl_base
```

We start to define the format of the output. We highlight items that are
not discussed in the previous discussion.

The `text_indent_first` and `test_indent_left` arguments are used to
control the indent space of the text. They are helpful when you need to
control the white space of a long sentence. For example, "AMERICAN
INDIAN OR ALASKA NATIVE" in the table.

```{r}
colheader1 <- paste(names(tbl_base), collapse = "|")
colheader2 <- paste(tbl_base[1, ], collapse = "|")
rel_width <- c(2.5, rep(1, 4))

tbl_base[-1, ] %>% 
      rtf_title("Participant Baseline Characteristics", 
                "(All Participants Randomized)") %>%
      rtf_colheader(colheader1, 
                    col_rel_width = rel_width) %>% 
      rtf_colheader(colheader2, 
                    border_top = "",
                    col_rel_width = rel_width) %>%
      rtf_body(col_rel_width = rel_width,
               text_justification = c("l", rep("c", 4)), 
               text_indent_first = -240,
               text_indent_left = 180) %>% 
      rtf_encode() %>% 
      write_rtf("tlf/tlf_base.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/tlf_base.pdf")
```

The procedure to generate a baseline characteristics table can be
summarized as follow:

-   Step 1: Read the data set.
-   Step 2: Use `table1::table1()` to get the baseline characteristics
    table.
-   Step 3: Transfer the output in Step 2 into a data frame that only
    contains ASCII characters.
-   Step 4: Define the format of the rtf table by using R package
    `r2rtf`.