### 2021-10-07 slides/workshop-slides.Rmd

---
title: "R for Clinical Study Reports and Submission"
subtitle: "R/Pharma Conference 2021"
author: "Yilong Zhang, Nan Xiao, Keaven Anderson"
date: "October, 2021"
output:
  mkdocs::mk_ioslides_presentation:
    mk_logo: "invent"
    mk_col_title: "white"
    mk_col_section: "teal"
    mk_class: "public"
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, eval = FALSE, include = FALSE}
rmarkdown::render("slides/workshop-slides.Rmd")
```

# Welcome

## Outline

In this workshop, we have three parts:

- Delivering TLFs in CSR (Yilong Zhang)
  + Session 1 (45 min)
  + Break (10 min)
  + Session 2 (45 min)
  + Break (20 min) 
- Clinical trial project (Nan Xiao)
  + Session 3 (40 min)
  + Break (10 min)
- eCTD submission package (Nan Xiao)
  + Session 4 (30 min)
  + Q&A (10 min)

## Disclaimer

- All opinions expressed are those of the presenter and not 
  Merck Sharp & Dohme Corp., a subsidiary of Merck & Co., Inc., Kenilworth, NJ, USA.

- Some slides need to be scrolled down to see the full content.

## Preparation

In this workshop, we assume you have 
some R programming experience and clinical development knowledge.

- Data manipulation: `tidyverse`, `dplyr`, `tidyr`, `ggplot2` etc. 
- ADaM data: `adsl`, `adae` etc.

```{r, echo = TRUE, eval = FALSE}
install.packages(c(
 "tidyverse",        # Data Manipulation 
 "r2rtf",            # TLF generation
 "pkglite"           # eCTD Submission 
))
```

## Resource

- Training Material: https://r4csr.org/
- Demo project: https://github.com/elong0527/esubdemo
- R Consortium FDA submission pilot: https://github.com/RConsortium/submissions-pilot1

## Philosophy

We share the same philosophy described in [Section 1.1 of R Packages book](https://r-pkgs.org/intro.html?q=Marwick#intro-phil) and quote here.

- "Anything that can be automated, should be automated." 
- "Do as little as possible by hand. Do as much as possible with functions." 
- "The goal is to spend your time thinking about what you want to do 
rather than thinking about the minutiae of package structure."

# Delivering TLFs in CSR

## ICH E3 guidance

The [ICH E3: strcuture and content of clinical study reports](https://database.ich.org/sites/default/files/E3_Guideline.pdf) provide guidance to assist sponsors in the development of a CSR.

In a CSR, most of TLFs are located in

- Section 10: Study patients
- Section 11: Efficacy evaluation
- Section 12: Safety evaluation
- Section 14: Tables, Figures and Graphs referred to but not included in the text
- Section 16: Appendices

## Datasets

- Public available CDISC pilot [study data located at CDISC Bitbucket repository](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse/updated-pilot-submission-package/900172/m5/datasets/cdiscpilot01/analysis/adam/datasets).

- The dataset structure is following [CDISC Analysis Data Model (ADaM)](https://www.cdisc.org/standards/foundational/adam).

- Source data: https://github.com/elong0527/r4csr/tree/master/adam_data

## Tools

- `tidyverse`: a collection of R packages to simplify the workflow to manipulate, 
visualize and analyze data in R.

- `r2rtf`: an R package to create production ready tables and figures in RTF format.

## r2rtf Motivation

In the pharmaceutical industry, RTF/Microsoft Word play a central role 
in preparing clinical study reports

Different organization can have different table standards

- E.g., Table layout, Font size, Border type, Footnote, Data source

`r2rtf` is designed to:

- Generate highly customized tables
- Limit package dependency
- Target regulatory deliverables   
- Support pipes (`%>%`)

- `r2rtf` Introduction: https://r4csr.org/overview.html#r2rtf

## Disposition Table

https://r4csr.org/disposition.html

## Analysis Population

https://r4csr.org/population.html

# Break (10 min)

## Baseline Characteristics

https://r4csr.org/baseline-characteristics.html

## Efficacy Table

https://r4csr.org/efficacy.html

## AE Summary Table

https://r4csr.org/aesummary.html

## Specific AE Table

https://r4csr.org/specific-ae.html

# Clinical Trial Project

# Break (20 min)

## Overview

In a late-stage clinical trial, the number of A&R deliverables can easily be in the hundreds.

For an organization, it is also common to have multiple ongoing clinical trials in a clinical program.

In this part, let’s consider how to organize a clinical trial project as an A&R leader.

## Folder Structure

https://r4csr.org/folder.html

## Project Management

https://r4csr.org/manage.html

# Break (10 min)

# eCTD Submission

# Q&A