### 2021-10-09 project-folder.Rmd
minor fixes for ch10 ch11

# (PART) Clinical Trial Project {-}

# Overview

In a late-stage clinical trial, the number of A&R deliverables can easily be in the hundreds. 
For an organization, it is also common to have multiple ongoing clinical trials in a clinical program.

To deliver A&R results of a clinical trial project, it is teamwork that typically requires
collaborations from both statisticians and programmers. 
In this part, let's consider how to organize a clinical trial project as an A&R leader.

In Chapter \@ref(folder), we will discuss how to organize source code, documents, 
deliverables in an A&R clinical project. 
We recommend using the R package folder structure.

In Chapter \@ref(manage), we will discuss a process or system development lifecycle 
to manage the A&R of a clinical project. 
We recommend following an agile management approach to define, develop, validate and deliver work.

# Project Folder {#folder}

A clearly defined clinical project folder structure can have many benefits to 
clinical development teams in an organization. 
Specifically, a well-defined project structure can achieve:

- Consistency: everyone works on the same folder structure.
- Reproducibility: analysis can be executed and reproduced by different team members months/years later.  
- Automation: automatically check the integration of a project.  
- Compliance: reduce compliance issue

We illustrate the R package project folder for the A&R project using a demo project called 
[`esubdemo`](https://github.com/elong0527/esubdemo.git). 
You can [clone](https://happygitwithr.com/rstudio-git-github.html#clone-the-new-github-repository-to-your-computer-via-rstudio) the project using RStudio.

For R users, you already benefit from a well-defined and consistent folder structure. 
That is the [R package folder structure](https://github.com/rstudio/cheatsheets/raw/master/package-development.pdf). 
Every R package developer is required to follow the same convention to organize their R
functions before the R package can be disseminated through the Comprehensive R Archive Network
(CRAN). As a user, you can easily install and use those R packages after installing from CRAN. 
There are many good resources to guide developers on R package development.
For example, Hadley Wickham's [R Packages book](https://r-pkgs.org/).

We recommend using the R package folder structure to organize analysis scripts 
for clinical trial development. 
Using the R package folder structure to streamline data analysis work has also been proposed before 
(@marwick2018packaging, @wuanalysis).

## Consistency {#consistency}

For consistency, a well-defined folder structure with potential templates
ensures project teams organize A&R work consistently across multiple projects. 
Consistent folder structure also reduces communication costs between study team members 
and enhances the transparency of projects.

In this book, we refer to an R package as **project-specific R package** 
if the purpose of an R package is to organize analysis scripts for a clinical project.

We refer to an R package as a **standard R package** if the purpose of an R package is to 
share commonly used R functions to be hosted in a code repository such as CRAN.

Below is minimal sufficient folders and files for a project-specific R package
based on the R package folder structure.

- `*.Rproj`: RStudio project file used to open RStudio project.
- `DESCRIPTION`: Metadata for a package including authors, license, dependencies, etc.
- `R/`: Project-specific R functions.
- `vignettes/`: Analysis scripts using Rmarkdown.
- `man/`: Manual of project-specific R functions.

A general discussion of the R package folder structure can be found 
in Chapter 4 of the R package book (@wickham2015r).

We demonstrate the idea using the [`esubdemo`](https://github.com/elong0527/esubdemo) project.

In the `esubdemo` project, we saved all TLF generation scripts in previous chapters into the `vignettes/` folder.

> Here, the `adam/` folder contains ADaM datasets. 
> The `tlf` folder contains output TLFs in RTF format.
> We put `adam/` and `tlf/` folders within the `vignettes` folder only for illustration purposes. 
> In an actual A&R report, you may have a different location to save your input and output.

```
vignettes
├── adam_data
├── tlf
├── tlf-01-disposition.Rmd
├── tlf-02-population.Rmd
├── tlf-03-baseline.Rmd
├── tlf-04-efficacy.Rmd
├── tlf-05-ae-summary.Rmd
└── tlf-06-ae-spec.Rmd
```

While creating those analysis scripts, we also defined a few helper functions (e.g. `fmt_num` and `count_by`). 
Those functions are saved in the `R/` folder.

```
R/
├── count_by.R
└── fmt.R
```

With a formal project, it is also important to provide proper documentation for those help functions. 
We used Roxygen2 to document functions. 
For example, the header below defines each variable in fmt_est. 
More details can be found in [Chapter 10 of the R package book](https://r-pkgs.org/man.html)

```{r, eval = FALSE}
#' Format point estimator
#'
#' @param .mean mean of an estimator.
#' @param .sd sd of an estimator.
#' @param digits number of digits for `.mean` and `.sd`.
#'
#' @export
fmt_est <- function(.mean,
                    .sd,
                    digits = c(1, 2)) {
  .mean <- fmt_num(.mean, digits[1], width = digits[1] + 4)
  .sd <- fmt_num(.sd, digits[2], width = digits[2] + 3)
  paste0(.mean, " (", .sd, ")")
}
```

Those `Roxygen2` documentation will be transferred to 
standard R documentation `.Rd` files saved in the `man/` folder. 
This step is automatically handled by `devtools::document()` function.

```
man
├── count_by.Rd
├── fmt_ci.Rd
├── fmt_est.Rd
├── fmt_num.Rd
└── fmt_pval.Rd
```

The `man` folder is used to save documentation automatically generated by `roxygen2`.
A typical workflow is to add `roxygen2` documentation at the header of each function in the `R/` folder. 
Then `devtools::documents()` is used to generate all the documentation files in the `man/` folder. 
More details can be found in [Chapter 10 of the R packages book](https://r-pkgs.org/man.html).

## Reproducibility {#reproduce}

Reproducibility of analysis is one of the most important aspects of regulatory deliverables. 
By using the R package folder structure and proper tools (e.g. `renv`, `packrat`), 
we illustrate how to achieve reproducibility for R and R package versions.

> This is the same level of reproducibility in most SAS environments.
> https://support.sas.com/techsup/pcn/altopsys.html

In the `esubdemo` project, a reproducible environment is created when you open the 
`esubdemo.Rproj` from RStudio. 
When we open the `esubdemo` project, RStudio will execute the R code in `.Rprofile` automatically.
So we can use `.Rprofile` to set up a reproducible environment. 
More details can be found in [https://rstats.wtf/r-startup.html](https://rstats.wtf/r-startup.html).

> `.Rprofile` is only for project-specific R package. 
> A standard R package should not use `.Rprofile`.

After we open the `esubdemo` project, the code in `.Rprofile` will automatically check 
the current R version is the same as we defined in `.Rprofile`

```{r, eval = FALSE}
# set up project R version
R_version <- "4.1.1"
```

If the R version mismatch, an error message is displayed as below.

```
Error: The current R version is not the same as the current project in R4.1.1
```

The R package version is controlled in two layers. 
Firstly, we define a snapshot date in `.Rprofile`. 
The snapshot date allows us to freeze the source code repository.

```{r, eval = FALSE}
# set up snapshot date
snapshot  <- "2021-08-06"

# set up repository based on the snapshot date
repos <- paste0("https://mran.microsoft.com/snapshot/", snapshot)

# define repo URL for project-specific package installation
options(repos = repos)
```

We can also define the package repository to be a specific snapshot date.
For example, we used Microsoft R Application Network (MRAN) to 
define the snapshot date to be `2021-08-06`. 
The snapshot date freezes the R package repository.

In other words, all R packages installed in this RStudio project 
are based on the frozen R version at the snapshot date. 
Here it is `2021-08-06` by using the MRAN server.

Below information will be displayed after a new R session is opened.

```
Current project R package repository:
    https://mran.microsoft.com/snapshot/2021-08-06
```

> RStudio Package Manager (RSPM) provided a solution to host both 
> publicly available and internally developed R packages. 
> However, the public RSPM server (https://packagemanager.rstudio.com/client/#/repos/2/overview) 
> did not provide a daily snapshot as MRAN had.

Secondly, we use `renv` to lock R package versions and save them in the `renv.lock` file.
`renv` provides a robust and stable approach to managing R package versions 
for project-specific R packages.
An introduction of `renv` can be found on its 
[website](https://rstudio.github.io/renv/articles/renv.html).

```{r, eval = FALSE}
source("renv/activate.R")
```

The R code above in the `.Rprofile` initiates the `renv` running environment.  
As a user, you can use `renv::init()`, `renv::snapshot()` and `renv::restore()` 
to initialize, save and restore R packages used for the current analysis project.

In the analysis project, the `renv` package will

- create a `renv.lock` file to save the state of package versions.
- create a `renv` folder to manage R packages for a project.

> `renv.lock` file and `renv/` folder are only for project-specific R package.
> A standard R package should not use `renv`.

In summary, the R package version is controlled in two layers.

- Define a snapshot date in `inst/startup.R`.
- Using `renv` to lock R versions within a project.

If the project is initiated properly  
you shall be able to see similar messages to inform how we control R package versions.

```
* Project '~/esubdemo' loaded. [renv 0.14.0]
```

Once R packages have been properly installed, 
the system will use the R packages located in the search 
path defined based on the order of `.libPaths()`.
The startup message also provided the R package search path.

```
Below R package path are searching in order to find installed R pacakges in this R session:
    "/home/zhanyilo/github-repo/esubdemo/renv/library/R-4.1/x86_64-pc-linux-gnu"
    "/rtmp/RtmpT3ljoY/renv-system-library"   
```

> A cloud-based R environment (e.g., RStudio Workbench)
> can enhance the reproducibility within an organization
> by using the same operating system, R version, and R package versions for an A&R project.
> More details can be found in [https://environments.rstudio.com/](https://environments.rstudio.com/).

> A container solution like [Docker](https://github.com/rstudio/r-docker) could further enhance 
> the reproducibility across an organization at the operating system level, but beyond the scope of this book.

In conclusion, to achieve reproducibility for a project-specific R package, 
a clinical project team can work under a controlled R environment 
in the same R version and R package versions defined by a repository snapshot date.

## Automation

By using the R package folder structure, you will benefit from many outstanding tools to 
simplify and streamline your workflow.

We have learned a few functions in `devtools` to generate content automatically.
Here is a shortlist of tools that can enhance the workflow.

- [`devtools`](https://devtools.r-lib.org/): make package development easier. 
  + A good overview can be found in [Chapter 2 of the R packages](https://r-pkgs.org/whole-game.html) book  
  + `devttols::load_all()`: load all functions in `R/` folder and running environment. 
  + `devtools::document()`: automatically create documentation using `roxygen2`.
  + `devttols::check()`: automatically perform compliance check as an R package. 
  + `devtools::build_site()`: automatically run analysis scripts in batch and create a `pkgdown` website. 
- [`usethis`](https://usethis.r-lib.org/): automates repetitive tasks that arise during project setup and development.
- [`testthat`](https://testthat.r-lib.org/): streamline testing code.
  + A discussion of using the `testthat` for an A&R project can be found in (@madhutest).  
- [`pkgdown`](https://pkgdown.r-lib.org/): generate static HTML documentation website for an R package
  + It also allows you to run all analysis code in batch.

You may further automatically execute routines by leveraging CI/CD workflow. 
For example, the `esubdemo` project will rerun all required checks and build a pkgdown website
by using [Github Actions](https://usethis.r-lib.org/reference/github_actions.html).

As the consistent folder is defined, it also becomes easier to create specific tools 
that fit for the analysis and reporting purpose.

Below are few potential tools that can be helpful:

- Create project template using [RStudio Project Templates](https://rstudio.github.io/rstudio-extensions/rstudio_project_templates.html)
- Add additional compliance checks for analysis and reporting. 
- Save log files for running in batch.

## Compliance

For a regulatory deliverable, it is important to maintain compliance. 
With a consistent folder structure, we can define specific criteria for compliance. 
Some compliance criteria can be implemented within the automatically checking steps.

For an R package, there are already criteria to ensure R package integrity.  
More details can be found in [Chapter 19 of the R package book](https://r-pkgs.org/r-cmd-check.html).