### 2021-10-12 submission-environment.Rmd

# Running Environment {#running-environment}

In the previous chapter, we have generated instructions to manually
create the running environments for reproducing the A&R deliverables.

In this chapter, we will focus on automating the creation of
the R environments with R code, to accelerate the dry run testing process,
simplify the ADRG instructions, and make it easy to recreate
different environment settings with reproducible analysis results.

## Prerequisites

`cleanslate` is an R package that offers a solution to
create portable R environments.

::: {.rmdnote}
As of Q4 2021, the cleanslate package used in this chapter
is still under active development and validation.
This chapter gives a preview of the planned APIs.
They may change in the future.
:::

Install `cleanslate` from CRAN (once available):

```{r, eval=FALSE}
install.packages("cleanslate")
```

Or from GitHub (once available):

```{r, eval=FALSE}
remotes::install_github("Merck/cleanslate")
```

## Practical Considerations

The `cleanslate` package supports:

- Creating a project folder with project-specific context
  (`.Rproj`, `.Rprofile`, `.Renviron`)
- Installing a specific version of R into the project folder
- Installing a specific version of Rtools into the project folder

An essential feature of `cleanslate` is that it does **not** require
administrator privileges to run R and Rtools installers.
This makes it easier to deploy under enterprise settings
and avoids security and portability concerns.

As many of the A&R deliverables are currently created, validated, and delivered
under Windows, the primary focus is Windows at the moment,
while the support for other platforms might be added in future versions.

## Create Canonical Environments

One can create a running environment with "canonical" settings with a
single function call to `use_cleanslate()`:

```{r, eval=FALSE}
cleanslate::use_cleanslate(
  "C:/temp/",
  r_version = "4.1.1",
  from = "https://cran.r-project.org/",
  repo = "https://mran.microsoft.com/snapshot/2021-08-06/"
)
```

This will

- Create a project folder under `C:/temp/` with a `.Rproj` file;
- Download R 4.1.1 installer from CRAN, install it into `C:/temp/R/`;
- Not install Rtools (by default, `rtools_version = NULL`);
- Create a `.Rprofile` file under the project folder, set `options(repos)`
  to use the specified `repo` (an MRAN snapshot in this example),
  and instructions to set the R binary path in RStudio;
- Create a `.Renviron` file under the project folder and set the
  library path to be the library of the project-specific R installation.

As a principle, one should always double-click the `.Rproj` file to
open the project.
This will ensure some sanity checks in the `.Rprofile`, such as whether the
R and library are located within the project folder.

## Create Tailored Environments

To create a more customized running environment, one can use the specific
functions to tailor each aspect, for example:

```{r, eval=FALSE}
library("cleanslate")

"C:/temp/" %>%
  use_project() %>%
  use_rprofile() %>%
  use_renviron() %>%
  use_r_version(version = "4.1.1") %>%
  use_rtools(version = "rtools40")
```

The project context functions
(`use_project()`, `use_rprofile()`, `use_renviron`) support custom templates
using [brew](https://cran.r-project.org/package=brew).

The `use_r_*()` functions have variations that serve as shortcuts
to use R versions defined by release lifecycles, for example,
`use_r_release()`, `use_r_oldrel()`, and `use_r_devel()`.
Note that to ensure better reproducibility, one should still use
`use_r_version()` as the release, oldrel, and devel versions will
shift as time goes by.

The helper functions `version_*()` and `snapshot_*()` could assist you
to determine specific versions of R and Rtools that are currently available,
besides generating and verifying the snapshot repo links.

## Update ADRG

If using cleanslate, remember to update the ADRG instructions for
executing the analysis programs in R. Mostly, this would simplify
the first three steps on creating a project, installing a specific
version of R, and configuring the package repo location.

## RStudio Addin

To make it convenient to use `cleanslate` in experiments,
one can also use its RStudio addin. After `cleanslate`
is installed, click `Addins` -> `cleanslate` -> `Create portable R environment`
in RStudio, or call `cleanslate:::create_env_addin()` to open it.

```{r, out.width="65%", echo=FALSE, fig.cap="cleanslate RStudio addin"}
knitr::include_graphics("images/cleanslate-addin.png")
```

The addin provides a wizard-like interface to help create the environment
with the most important options, although with less flexibility compared
to the functional API demonstrated above.