### 2021-10-14 index.Rmd
use sentence case instead of title case for headings; run styler on Rmds

--- 
title: "R for Clinical Study Reports and Submission"
author: ["Yilong Zhang", "Nan Xiao", "Keaven Anderson"]
knit: "bookdown::render_book"
description: >
  This book will teach you how to prepare tables, listings, and figures for
  clinical study report and submit to regulatory agencies, 
  the essential part of clinical trial development.

site: bookdown::bookdown_site

documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
---

```{r, include = FALSE}
library(bslib)
library(xml2)
library(downlit)
library(haven)
library(dplyr)
library(tidyr)
library(r2rtf)
library(knitr)
library(kableExtra)
library(pkglite)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
options(dplyr.summarise.inform = FALSE)
file.copy("static/CNAME", "_book", overwrite = TRUE)
```

```{r, eval = FALSE, include=FALSE}
r2rtf:::rtf_convert_format(
  list.files("tlf", pattern = ".rtf", full.names = TRUE),
  output_dir = "tlf"
)
```

```{r, eval = FALSE, include=FALSE}
rsconnect::writeManifest(appPrimaryDoc = "index.Rmd", contentCategory = "site")
```

```{r, eval = FALSE, include=FALSE}
bookdown::render_book("index.Rmd", "bookdown::pdf_book")
```

# Welcome! {-}

<img src="images/cover.jpg" width="100%" align="right" alt="" class="cover" />

Welcome to R for clinical study reports and submission. 
Clinical study reports (CSR) are essential parts of clinical trial development. 
A CSR is an "integrated" full scientific report of an individual clinical trial.

The [ICH E3: structure and content of clinical study reports](https://database.ich.org/sites/default/files/E3_Guideline.pdf) 
provides guidance to assist sponsors in the development of a CSR. 
In this book, you will learn how to use R to prepare a CSR and
how to submit it to regulatory agencies.

**This is the work-in-progress draft of the book.**

## Authors and contributors {-}

```{r, echo = FALSE}
contributors <- c(
  person("Yilong", "Zhang", email = "yilong.zhang@merck.com", role = "aut"),
  person("Nan", "Xiao", role = "aut"),
  person("Keaven", "Anderson", role = "aut"),
  person("Yujie", "Zhao", role = "ctr"),
  person("Aiming", "Yang", role = "ctr"),
  person("Steven", "Haesendonckx", role = "ctr")
)
```

The document is maintained by a community.
While reading the document, you can be a contributor as well.
The quality of this document relies on you.

- Author: contribute the majority content of at least one chapter.
- Contributor: [contribute at least one commit to the source code](https://github.com/elong0527/r4csr)

```{r, echo = FALSE, eval = TRUE}
data.frame(contributors = as.character(contributors)) %>%
  kable(format = "html") %>%
  kable_styling()
```