### 2021-10-14 intro.Rmd

# Introduction {#intro}

## Folder structure

In clinical trial development, source code needs to be developed and maintained to generate and deliver 
Study Data Tabulation Model (SDTM), Analysis Dataset Model (ADaM) datasets and tables, listings, and figures (TLFs).
A typical example is a Phase 3 clinical trial where hundreds of TLFs are required for submission. 
Considering the amount of programs required for such an effort, 
a consistent and well-defined folder structure is crucial in 
managing a clinical trial analysis and reporting (A&R) project.

We recommend using 
[R package folder structure](https://github.com/rstudio/cheatsheets/raw/master/package-development.pdf) 
to organize all the A&R related source code and documentation for a clinical trial A&R project. 
The R package folder structure is well defined and widely used in R community through repositories (e.g. CRAN).

The consistent approach of using R package folder structure 
simplifies the communication for all developers within and across organizations.

- For a new R developer, 
it is an essential step to develop R packages 
when you want to share your work with others. 
You will learn one folder structure that is widely used 
in the R community with outstanding tutorials and tools for free.

- For an experienced R developer, there is a minimal learning curve.

- For an organization, it simplifies the process, tool, template, 
and training development, because a unified folder structure is used 
to develop and maintain standard tool and analysis projects.

The workflow around an R package can also improve 
the traceability and reproducibility for an analysis project [@marwick2018packaging].

We will revisit folder structure when we discuss project management for a clinical trail project.

In addition, R package folder structure is also recommended 
to develop Shiny apps as discussed 
in [Chapter 20 of Master Shiny](https://mastering-shiny.org/scaling-packaging.html) book 
and [Engineering Production-Grade Shiny Apps](https://engineering-shiny.org/golem.html) book.

## In this book

This book is an intermediate-level book by 
assuming readers have R programming and clinical development knowledge.
The assumption of each part is as below:

- Part 1, "Delivering TLFs in CSR" provides general information with examples 
to create tables, listings, and figures. 
In this part, we assume readers are individual contributors to a clinical project
with some experience in R programming. 
We expect readers are familiar with data manipulation in R. 
Some good references include [Hands-On Programming with R](https://rstudio-education.github.io/hopr/),
[R for Data Science](https://r4ds.had.co.nz/) and 
[Data Manipulation with R](https://www.amazon.com/Data-Manipulation-R-Use/dp/0387747303).

- Part 2, "Clinical Trial Project" provides general information 
with examples to manage a clinical trial A&R project. 
In this part, we assume a reader is a project lead with experience in R package development. 
Some good reference includes [R packages](https://r-pkgs.org/) and 
[The tidyverse style guide](https://style.tidyverse.org/).

- Part 3, "eCTD submission package"
provides general information in preparing 
submission package related to clinical study report (CSR) in 
electronic Common Technical Document (eCTD). 
In this part, we assume a reader is a project lead of a clinical project 
with experience in R package development and submission.

## Philosophy

We share the same philosophy described in 
[Section 1.1 of R Packages book](https://r-pkgs.org/intro.html#intro-phil) 
and quote here.

- "Anything that can be automated, should be automated." 
- "Do as little as possible by hand. Do as much as possible with functions." 
- "The goal is to spend your time thinking about what you want to do 
rather than thinking about the minutiae of the package structure."