### 2021-10-27 slides/workshop-slides.Rmd
reorg r2rtf examples

---
title: "R for Clinical Study Reports and Submission"
subtitle: "R/Pharma Conference 2021"
author: "Yilong Zhang, Nan Xiao, Keaven Anderson"
date: "October, 2021"
output:
  mkdocs::mk_ioslides_presentation:
    mk_logo: "invent"
    mk_col_title: "white"
    mk_col_section: "teal"
    mk_class: "public"
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(dplyr)
library(tidyr)
library(r2rtf)
```

```{r, eval = FALSE, include = FALSE}
rmarkdown::render("workshop-slides.Rmd")
```

# Welcome

## Outline

In this workshop, we have three parts:

- Delivering TLFs in CSR (Yilong Zhang)
  + Session 1 (45 min)
  + Break (10 min)
  + Session 2 (45 min)
  + Break (20 min) 
- Clinical trial project (Nan Xiao)
  + Session 3 (30 min)
  + Break (10 min)
- eCTD submission package (Nan Xiao)
  + Session 4 (40 min)
  + Q&A (10 min)

## Disclaimer

- All opinions expressed are those of the presenter and not 
  Merck Sharp & Dohme Corp., a subsidiary of Merck & Co., Inc., Kenilworth, NJ, USA.

- Some slides need to be scrolled down to see the full content.

## Training Objective

- Learning how to create tables for clinical study reports.
- Learning how to organize a clinical development project.
- Learning how to prepare eCTD submission package to FDA.

> The tool, process and format may be different in different organization/company.
> We only provide a recommended way to address them.

## Acknowledgement

- Team members from Merck & Co. 
  + [We are hiring!](https://jobs.merck.com/bards) 
  + Contact us if interested.

- [Contributors](https://r4csr.org/) of the training materials.
  + Please consider to submit issues or PR in the repo.

- [R consortium](https://www.r-consortium.org/projects/isc-working-groups)
  + R validation hub
  + Submission working group
  + R Tables for Regulatory Submission (RTRS) working group
  + Please consider to join and contribute to a working group.

- R / Pharma organizers
  + It is a fun and productive annual gathering. 
  + Please consider to share story and use cases to enlarge the community.

## Preparation

In this workshop, we assume you have 
some R programming experience and clinical development knowledge.

- Data manipulation: `tidyverse`, `dplyr`, `tidyr`, `ggplot2` etc. 
- ADaM data: `adsl`, `adae` etc.

```{r, echo = TRUE, eval = FALSE}
install.packages(c(
 "tidyverse",        # Data Manipulation 
 "r2rtf",            # TLF generation
 "pkglite"           # eCTD Submission 
))
```

## Resource

- Training Material: https://r4csr.org/
- Demo project: 
  + Analysis project: https://github.com/elong0527/esubdemo
  + eCTD submission package: https://github.com/elong0527/ectddemo
- R Consortium FDA submission pilot: 
  + https://github.com/RConsortium/submissions-pilot1

- During the workshop, we will use `esubdemo` project.
  + `esubdemo` is shared in RStudio Cloud (https://rstudio.cloud/) 
  + Project link will be shared in chat
  + Post questions in group chat.

## Philosophy

We share the same philosophy described in 
[Section 1.1 of R Packages book](https://r-pkgs.org/intro.html?q=Marwick#intro-phil) and quote here.

- "Anything that can be automated, should be automated." 
- "Do as little as possible by hand. Do as much as possible with functions." 
- "The goal is to spend your time thinking about what you want to do 
rather than thinking about the minutiae of package structure."

# Delivering TLFs in CSR

## ICH E3 guidance

The [ICH E3: strcuture and content of clinical study reports](https://database.ich.org/sites/default/files/E3_Guideline.pdf) provide guidance to assist sponsors in the development of a CSR.

In a CSR, most of TLFs are located in

- Section 10: Study patients
- Section 11: Efficacy evaluation
- Section 12: Safety evaluation
- Section 14: Tables, Figures and Graphs referred to but not included in the text
- Section 16: Appendices

## Datasets

- Public available CDISC pilot [study data located at CDISC Bitbucket repository](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse/updated-pilot-submission-package/900172/m5/datasets/cdiscpilot01/analysis/adam/datasets).

- The dataset structure is following [CDISC Analysis Data Model (ADaM)](https://www.cdisc.org/standards/foundational/adam).

- Source data: https://github.com/elong0527/r4csr/tree/master/adam_data

## Tools

- `tidyverse`: a collection of R packages to simplify the workflow to manipulate, 
visualize and analyze data in R.

- `r2rtf`: an R package to create production ready tables and figures in RTF format.

# r2rtf Introduction

## Motivation

In the pharmaceutical industry, RTF/Microsoft Word play a central role 
in preparing clinical study reports

- [ICH E3 Structure and Content of Clinical Study Reports](https://database.ich.org/sites/default/files/E3_Guideline.pdf)

Different organization can have different table standards

- E.g., Table layout, Font size, Border type, Footnote, Data source

- `r2rtf` is an R package to create production-ready tables and figures in RTF format.
The R package is designed to

`r2rtf` is designed to:

- provide simple "verb" functions that correspond to each component of a table, 
  to help you translate a data frame to a table in RTF file.
- enable pipes (`%>%`). 
- only focus on the **table format**. 
  Data manipulation and analysis shall be handled by other R packages. (e.g. `tidyverse`)

## Workflow

Before creating an RTF table we need to:

- Figure out table layout.

- Split the layout into small tasks in the form of a computer program.

- Execute the program.

## Minimal Example

`r2rtf` is designed to enable pipes (`%>%`)

```{r, echo = TRUE, eval = FALSE}
head(iris) %>%
  rtf_body() %>%                     # Step 1 Add table attributes 
  rtf_encode() %>%                   # Step 2 Convert attributes to RTF encode 
  write_rtf("minimal.rtf")           # Step 3 Write to a .rtf file
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("minimal.pdf")
```

## Package Overview

`r2rtf` package provides the flexibility to customize table appearance for

- **Table component**: title, column header, footnote, etc.
- **Table cell style**: size, border type, color, font size, text color, alignment, etc.
- **Flexible control**: the specification of the cell style can be row or column vectorized.
- **Complicated format**: pagination, section grouping, multiple table concatenations, etc.

`r2rtf` package also provides the flexibility to convert figures in RTF format.

## Simple Example - Adverse Events

`r2rtf` only focuses on **table format**. 
Data manipulation and analysis shall be handled by other R packages. (e.g. `tidyverse`)

- Let's refer examples in the [Section 2.4.2](https://r4csr.org/overview.html#r2rtf)

## Function Summary

`r2rtf` provides simple "verb" functions that correspond to each component of a table, 
  to help you translate data frame to table in RTF file.

```{r, echo = FALSE}
r2rtf_funs <- read.csv("r2rtf-functions.csv")
names(r2rtf_funs)[3] <- "Optional/required"
knitr::kable(r2rtf_funs)
```

## Function Illustration

```{r, out.width="80%", echo = FALSE}
knitr::include_graphics("function_summary.PNG")
```

# Break (10 min)

# CSR Examples

## Disposition Table

https://r4csr.org/disposition.html

## Analysis Population

https://r4csr.org/population.html

## Baseline Characteristics

https://r4csr.org/baseline-characteristics.html

## Efficacy Table

https://r4csr.org/efficacy.html

## AE Summary Table

https://r4csr.org/aesummary.html

## Specific AE Table

https://r4csr.org/specific-ae.html

# Clinical Trial Project

# Break (20 min)

## Overview

In a late-stage clinical trial, the number of A&R deliverables can easily be in the hundreds.

For an organization, it is also common to have multiple ongoing clinical trials in a clinical program.

In this part, let’s consider how to organize a clinical trial project as an A&R leader.

## Folder Structure

https://r4csr.org/folder.html

## Project Management

https://r4csr.org/manage.html

# Break (10 min)

# eCTD Submission

# Q&A