### 2021-10-27 slides/workshop-slides.Rmd
add slides for ch 9 - 14

---
title: "R for Clinical Study Reports and Submission"
subtitle: "R/Pharma Conference 2021"
author: "Yilong Zhang, Nan Xiao, Keaven Anderson"
date: "October 2021"
output:
  mkdocs::mk_ioslides_presentation:
    mk_logo: "invent"
    mk_col_title: "white"
    mk_col_section: "teal"
    mk_class: "public"
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(dplyr)
library(tidyr)
library(r2rtf)
```

```{r, eval = FALSE, include = FALSE}
rmarkdown::render("workshop-slides.Rmd")
```

# Welcome

## Outline

In this workshop, we have three parts:

- Delivering TLFs in CSR (Yilong Zhang)
  + Session 1 (45 min)
  + Break and/or Exercise (20 min)
  + Session 2 (45 min)
  + Break (10 min) 
- Clinical trial project (Nan Xiao)
  + Session 3 (30 min)
  + Break (10 min)
- eCTD submission package (Nan Xiao)
  + Session 4 (40 min)
  + Q&A (10 min)

## Disclaimer

- All opinions expressed are those of the presenter and not 
  Merck Sharp & Dohme Corp., a subsidiary of Merck & Co., Inc., Kenilworth, NJ, USA.

- Some slides need to be scrolled down to see the full content.

## Training Objective

- Learning how to create tables for clinical study reports.
- Learning how to organize a clinical development project.
- Learning how to prepare eCTD submission package to FDA.

> The toolchain, process, and formats may be different in different organizations.
> We only provide a recommended way to address them.

## Acknowledgement

- Team members from Merck & Co. 
  + [We are hiring!](https://jobs.merck.com/bards) 
  + Contact us if interested.

- [Contributors](https://r4csr.org/) of the training materials.
  + Please consider submitting issues or PR in the repo.

- [R Consortium](https://www.r-consortium.org/projects/isc-working-groups)
  + R validation hub
  + Submission working group
  + R Tables for Regulatory Submission (RTRS) working group
  + Please consider to join and contribute to a working group.

- R / Pharma organizers
  + It is a fun and productive annual gathering.
  + Please consider sharing stories and use cases to expand the community.

## Preparation

In this workshop, we assume you have 
some R programming experience and clinical development knowledge.

- Data manipulation: `tidyverse`, `dplyr`, `tidyr`, `ggplot2` etc. 
- ADaM data: `adsl`, `adae` etc.

```{r, echo = TRUE, eval = FALSE}
install.packages(c(
  "tidyverse",      # Data manipulation
  "r2rtf",          # TLF generation
  "pkglite"         # eCTD submission
))
```

## Resource

- Training Material: <https://r4csr.org/>
- Demo project: 
  + Analysis project: <https://github.com/elong0527/esubdemo>
  + eCTD submission package: <https://github.com/elong0527/ectddemo>
- R Consortium FDA submission pilot: 
  + <https://github.com/RConsortium/submissions-pilot1>

- During the workshop, we will use `esubdemo` project.
  + `esubdemo` is shared in RStudio Cloud (<https://rstudio.cloud/>)
  + Project link will be shared in chat
  + Post questions in group chat.

## Philosophy

We share the same philosophy described in 
[Section 1.1 of R Packages book](https://r-pkgs.org/intro.html?q=Marwick#intro-phil) and quote here.

- "Anything that can be automated, should be automated." 
- "Do as little as possible by hand. Do as much as possible with functions." 
- "The goal is to spend your time thinking about what you want to do 
rather than thinking about the minutiae of package structure."

# Delivering TLFs in CSR

## ICH E3 Guidance

The [ICH E3: structure and content of clinical study reports](https://database.ich.org/sites/default/files/E3_Guideline.pdf) provide guidance to assist sponsors in the development of a CSR.

In a CSR, most of TLFs are located in

- Section 10: Study patients
- Section 11: Efficacy evaluation
- Section 12: Safety evaluation
- Section 14: Tables, Figures and Graphs referred to but not included in the text
- Section 16: Appendices

## Datasets

- Public available CDISC pilot [study data located at CDISC Bitbucket repository](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse/updated-pilot-submission-package/900172/m5/datasets/cdiscpilot01/analysis/adam/datasets).

- The dataset structure follows the [CDISC Analysis Data Model (ADaM)](https://www.cdisc.org/standards/foundational/adam).

- Source data: <https://github.com/elong0527/r4csr/tree/master/adam_data>

## Tools

- `tidyverse`: a collection of R packages to simplify the workflow to manipulate, 
visualize and analyze data in R.

- `r2rtf`: an R package to create production-ready tables and figures in RTF format.

# r2rtf Introduction

## Motivation

In the pharmaceutical industry, RTF/Microsoft Word play a central role 
in preparing clinical study reports

- [ICH E3 Structure and Content of Clinical Study Reports](https://database.ich.org/sites/default/files/E3_Guideline.pdf)

Different organizations can have different table standards

- E.g., Table layout, Font size, Border type, Footnote, Data source

- `r2rtf` is an R package to create production-ready tables and figures in RTF format.
The R package is designed to

`r2rtf` is designed to:

- provide simple "verb" functions that correspond to each component of a table, 
  to help you translate a data frame to a table in RTF file.
- enable pipes (`%>%`). 
- only focus on the **table format**. 
  Data manipulation and analysis shall be handled by other R packages. (e.g. `tidyverse`)

## Workflow

Before creating an RTF table, we need to:

- Figure out table layout.

- Split the layout into small tasks in the form of a computer program.

- Execute the program.

## Minimal Example

`r2rtf` is designed to enable pipes (`%>%`)

```{r, echo = TRUE, eval = FALSE}
head(iris) %>%
  rtf_body() %>%           # Step 1 Add table attributes
  rtf_encode() %>%         # Step 2 Convert attributes to RTF encode
  write_rtf("minimal.rtf") # Step 3 Write to a .rtf file
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("minimal.pdf")
```

## Package Overview

`r2rtf` package provides the flexibility to customize table appearance for

- **Table component**: title, column header, footnote, etc.
- **Table cell style**: size, border type, color, font size, text color, alignment, etc.
- **Flexible control**: the specification of the cell style can be row or column vectorized.
- **Complicated format**: pagination, section grouping, multiple table concatenations, etc.

`r2rtf` package also provides the flexibility to convert figures in RTF format.

## Simple Example - Adverse Events

`r2rtf` only focuses on **table format**. 
Data manipulation and analysis shall be handled by other R packages. (e.g. `tidyverse`)

- Let's refer to examples in the [Section 2.4.2](https://r4csr.org/overview.html#r2rtf)

## Function Summary

`r2rtf` provides simple "verb" functions that correspond to each component of a table, 
  to help you translate data frame to tables in RTF file.

```{r, echo = FALSE}
r2rtf_funs <- read.csv("r2rtf-functions.csv")
names(r2rtf_funs)[3] <- "Optional/required"
knitr::kable(r2rtf_funs)
```

## Function Illustration

```{r, out.width="80%", echo = FALSE}
knitr::include_graphics("function_summary.PNG")
```

# Break and/or Exercise (20 min)

# CSR Examples

## Disposition Table

<https://r4csr.org/disposition.html>

## Analysis Population

<https://r4csr.org/population.html>

## Baseline Characteristics

<https://r4csr.org/baseline-characteristics.html>

## Efficacy Table

<https://r4csr.org/efficacy.html>

## AE Summary Table

<https://r4csr.org/aesummary.html>

## Specific AE Table

<https://r4csr.org/specific-ae.html>

# Break (10 min)

# Clinical Trial Project

## Overview

In a late-stage clinical trial, the number of A&R deliverables can easily be in the hundreds.

For an organization, it is also common to have multiple ongoing clinical trials in a clinical program.

In this part, let's consider how to organize a clinical trial project as an A&R lead.

## Folder Structure

Our primary focus is creating a standard R package structure to organize the
project, with 4 goals in mind:

- Consistency
- Reproducibility
- Automation
- Compliance

<https://r4csr.org/folder.html>

## Project Management

- Setting up for success
  - Teamwork
  - Architecture
  - Boundaries
  - Community

- The SDLC
  - Planning
  - Development
  - Validation
  - Operation

<https://r4csr.org/manage.html>

# Break (10 min)

# eCTD Submission

## Overview

<https://r4csr.org/overview-2.html>

## Submission Package

We will discuss strategies to prepare proprietary R packages and analysis code
into proper formats for submission:

- The whole game
- Practical considerations
- Prepare R packages using `pkglite`
- Prepare analysis programs
- Update ADRG and ARM

<https://r4csr.org/submission-package.html>

## Running Environment

We will also give recommendations to make the R code
running environment reproducible for dry run tests and reviews
using `cleanslate`.

<https://r4csr.org/running-environment.html>

# pkglite

## pkglite: Compact Package Representations

- To provide a **tool** for packing and restoring R packages as plaintext
assets that are easy to store, transfer, and review.
- To provide a **grammar** for specifying the file packing scope that is
functional, precise, and extendable.
- To provide a **standard** for exchanging the packed asset that is
unambiguous, human-friendly, and machine-readable.

## Pipe-Friendly Workflow

```{r, eval=FALSE, echo=TRUE}
library("pkglite")

"/path/to/pkg/" %>%
  collate(file_ectd()) %>%
  pack()

pack(
  "/path/to/pkg1/" %>% collate(file_ectd()),
  "/path/to/pkg2/" %>% collate(file_ectd()),
  output = "/path/to/pkglite.txt"
)

"/path/to/pkglite.txt" %>%
  unpack(output = "/path/to/output/")
```

## File Specifications + File Collections

File specifications offer flexibility and brevity in specifying the files to include.

| File specification type | Functions                                                                           |
|-------------------------|-------------------------------------------------------------------------------------|
| Manual discovery        | `file_spec()`                                                                       |
| Automatic discovery     | `file_auto()`                                                                       |
| Common patterns         | `file_root_core()`, `file_r()`, `file_man()`, `file_src()`, `file_vignettes()`, ... |
| Default sets            | `file_default()`, `file_ectd()`                                                     |

<br>

File collections contain the evaluation results of file specifications for packing.

```{r, eval=FALSE, echo=TRUE}
library("pkglite")

"/path/to/pkg" %>%
  collate(file_root_core(), file_r(), file_auto("inst/"))
```

## pkglite.txt

pkglite.txt follows the standard Debian Control File (DCF) format used by
Debian, R, and RStudio IDE, to be both machine-readable and human-readable.

```
# Generated by pkglite: do not edit by hand
# Use pkglite::unpack() to restore the packages

Package: pkg1
File: DESCRIPTION
Format: text
Content:
  Package: pkg1
  Type: Package
  Title: Example Package One
  Version: 0.1.0
```

# Q&A