### 2021-11-08 project-overview.Rmd

# (PART) Clinical trial project {-}

# Overview {#project-overview}

In a late-stage clinical trial, the number of A&R deliverables can easily be in the hundreds.
For an organization, it is common to have multiple ongoing clinical trials in a clinical program.

To deliver the A&R results of a clinical trial project,
it is teamwork that typically requires collaborations from both statisticians and programmers.
In this part, let's consider how to organize a clinical trial project as an A&R lead.

Chapter \@ref(folder) will discuss how to organize source code, documents, and
deliverables in an A&R clinical project.
We recommend using the R package folder structure.

Chapter \@ref(manage) will discuss a process or system development lifecycle
to manage the A&R of a clinical project.
We recommend following an agile management approach to define, develop, validate, and deliver work.