### 2021-11-17 tlf-overview.Rmd

# (PART) Delivering TLFs in CSR {-}

# Overview {#tlf-overview}

## Background

In clinical trials, a critical step is to submit trial results to regulatory agencies.
[Electronic Common Technical Document (eCTD)](https://en.wikipedia.org/wiki/Electronic_common_technical_document)
has become a worldwide regulatory submission standard format.
For example, the United States Food and Drug Administration (US FDA) requires
new drug applications and biologics license applications
[must be submitted using the eCTD format](https://www.fda.gov/drugs/electronic-regulatory-submission-and-review/electronic-common-technical-document-ectd).
The Clinical Data Interchange Standards Consortium (CDISC) provides a
[pilot project following ICH E3 guidance](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse).

Within eCTD, clinical study reports (CSRs) are located at module 5.
[ICH E3 guidance](https://www.ich.org/page/efficacy-guidelines) provides
a compilation of the structure and content of clinical study reports.

A typical CSR contains full details on the methods and results of an individual clinical study.
In support of the statistical analysis, a large number of tables, listings, and figures are
incorporated into the main text and appendices.
In the CDISC pilot project, an
[example CSR](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse/updated-pilot-submission-package/900172/m5/53-clin-stud-rep/535-rep-effic-safety-stud/5351-stud-rep-contr/cdiscpilot01/cdiscpilot01.pdf)
is also provided. If you are interested in more examples of clinical study reports,
you can go to the [European Medicines Agency (EMA) clinical data website](https://clinicaldata.ema.europa.eu/web/cdp/home).

Building CSRs is teamwork between clinicians, medical writers, statisticians, statistical programmers,
and other relevant specialists such as experts on biomarkers.
Here, we focus on the work and deliverables completed by statisticians and statistical programmers.
In an organization, they commonly work together to
define, develop, validate and deliver tables, listings, and figures (TLFs) required for a CSR to
summarize the efficacy and/or safety of the pharmaceutical product.
Microsoft Word is widely used to prepare CSR in the pharmaceutical industry.
Therefore, `.rtf`, `.doc`, `.docx` are commonly used formats in their deliverables.

In this chapter, our focus is to illustrate how to create tables, listings, and figures (TLFs) in RTF format
that is commonly used in a CSR. The examples are in compliance with the
[FDA's Portable Document Format (PDF) Specifications](https://www.fda.gov/media/76797/download).

::: {.rmdnote}
FDA's PDF specification is a general reference. Each organization can define
more specific TLF format requirements that can be different from the examples in this book.
:::

## Structure and content

In the rest of this chapter, we are following the
[ICH E3 guidance](https://database.ich.org/sites/default/files/E3_Guideline.pdf)
on the structure and content of clinical study reports.

In a CSR, most of TLFs are located in

- Section 10: Study participants
- Section 11: Efficacy evaluation
- Section 12: Safety evaluation
- Section 14: Tables, listings, and figures referrals but not included in the text
- Section 16: Appendices

## Datasets

We used publicly available CDISC pilot
[study data located in the CDISC Bitbucket repository](https://bitbucket.cdisc.org/projects/CED/repos/sdtm-adam-pilot-project/browse/updated-pilot-submission-package/900172/m5/datasets/cdiscpilot01/analysis/adam/datasets).

For simplicity, we have downloaded all these datasets into the `data-adam/`
folder of this project and converted them from the `.xpt` format to
the `.sas7bdat` format.

The dataset structure follows
[CDISC Analysis Data Model (ADaM)](https://www.cdisc.org/standards/foundational/adam).

## Tools

In this part, we mainly use the R packages below to illustrate
how to deliver TLFs in a CSR.

- [tidyverse](https://www.tidyverse.org/): prepare datasets ready for reporting.
- [r2rtf](https://merck.github.io/r2rtf/): create RTF outputs

::: {.rmdnote}
There are other R packages to create TLFs in ASCIII, RTF and Word format.
For example, `rtables`, `huxtable`, `pharmaRTF`, `gt`, `officer`, `flextable` etc.
Here we focus on `r2rtf` to illustrate the concept.  
Reader is encourage to explore other R packages to find the proper tools to fit your purpose. 
:::

### tidyverse

`tidyverse` is a collection of R packages to simplify the workflow to manipulate,
visualize and analyze data in R.
Those R packages share
[the tidy tools manifesto](https://cran.r-project.org/web/packages/tidyverse/vignettes/manifesto.html)
and are easy to use for interactive data analysis.

RStudio provided outstanding [cheatsheets](https://www.rstudio.com/resources/cheatsheets/)
and [tutorials](https://github.com/rstudio-education/remaster-the-tidyverse) for `tidyverse`.

There are also books to introduce tidyverse.
We assume the reader to have experience in using `tidyverse` in this book.

- [The tidyverse cookbook](https://rstudio-education.github.io/tidyverse-cookbook/)
- [R for Data Science](https://r4ds.had.co.nz/)

### r2rtf

`r2rtf` is an R package to create production-ready tables and figures in RTF format.
This R package is designed to

- provide simple "verb" functions that correspond to each component of a table,
  to help you translate a data frame to a table in an RTF file;
- enable pipes (`%>%`);
- focus on the **table format** only.
  Data manipulation and analysis shall be handled by other R packages (e.g., `tidyverse`).

Before creating an RTF table, we need to

- figure out the table layout;
- split the layout into small tasks in the form of a computer program;
- execute the program.

We provide a brief introduction of `r2rtf` and showe how to transfer
data frames into table, listing, and figures (TLFs).

Other extended examples and features are covered on the
[`r2rtf` package website](https://merck.github.io/r2rtf/articles/index.html).

To explore the basic RTF generation verbs in `r2rtf`,
we will use the dataset `r2rtf_adae` saved in the `r2rtf` package.
This dataset contains adverse events (AEs) information from a clinical trial.

We will begin by loading the packages:

```{r}
library(dplyr) # Manipulate data
library(tidyr) # Manipulate data
library(r2rtf) # Reporting in RTF format
```

Below is the meaning of relevant variables.
More information can be found on the help page of the dataset (`?r2rtf_adae`)

In this example, we consider three variables:

- USUBJID: Unique Subject Identifier
- TRTA: Actual Treatment
- AEDECOD: Dictionary-Derived Term

```{r}
r2rtf_adae %>%
  select(USUBJID, TRTA, AEDECOD) %>%
  head(4)
```

`dplyr` and `tidyr` packages within `tidyverse` are used
for data manipulation to create a data frame
that contains all the information we want to add in an RTF table.

```{r}
tbl <- r2rtf_adae %>%
  count(TRTA, AEDECOD) %>%
  pivot_wider(names_from = TRTA, values_from = n, values_fill = 0)

tbl %>% head(4)
```

Now we have a dataset `tbl` in preparing the final RTF table.

`r2rtf` aims to provide one function for each type of table layout.
Commonly used verbs include:

- `rtf_page()`: RTF page information
- `rtf_title()`: RTF title information
- `rtf_colheader()`: RTF column header information
- `rtf_body()`: RTF table body information
- `rtf_footnote()`: RTF footnote information
- `rtf_source()`: RTF data source information

All these verbs are designed to enable the usage of pipes (`%>%`).
A full list of all functions can be found in the
[r2rtf package function reference manual](https://merck.github.io/r2rtf/reference/index.html).

A minimal example below illustrates how to combine verbs using pipes to create an RTF table.

- `rtf_body()` is used to define table body layout.
- `rtf_encode()` transfers table layout information into RTF syntax.
- `write_rtf()` save RTF encoding into a file with file extension `.rtf`

```{r}
head(tbl) %>%
  rtf_body() %>% # Step 1 Add table  attributes
  rtf_encode() %>% # Step 2 Convert attributes to RTF encode
  write_rtf("tlf/intro-ae1.rtf") # Step 3 Write to a .rtf file
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/intro-ae1.pdf")
```

If we want to adjust the width of each column to
provide more space to the first column,
this can be achieved by updating the `col_rel_width` argument
in the `rtf_body()` function.

In this example, the input of `col_rel_width` is a vector
with the same length for the number of columns.
This argument defines the relative width of each column
within a pre-defined total column width.

In this example, the defined relative width is `3:2:2:2`.
Only the ratio of `col_rel_width` is used.
Therefore it is equivalent to use `col_rel_width = c(6, 4, 4, 4)`
or `col_rel_width = c(1.5, 1, 1, 1)`.

```{r}
head(tbl) %>%
  rtf_body(col_rel_width = c(3, 2, 2, 2)) %>%
  # define relative width
  rtf_encode() %>%
  write_rtf("tlf/intro-ae2.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/intro-ae2.pdf")
```

In the previous example, we found the issue of a misaligned column header.
We can fix the issue by using the `rtf_colheader()` function.

In `rtf_colheader()`, the `colheader` argument is used to provide the content of the column header.
We use `"|"` to separate the columns.

In the example below, `"Adverse Events | Placebo | Xanomeline High Dose | Xanomeline Low Dose"`
define a column header with 4 columns.

```{r}
head(tbl) %>%
  rtf_colheader(
    colheader = "Adverse Events | Placebo | Xanomeline High Dose | Xanomeline Low Dose",
    col_rel_width = c(3, 2, 2, 2)
  ) %>%
  rtf_body(col_rel_width = c(3, 2, 2, 2)) %>%
  rtf_encode() %>%
  write_rtf("tlf/intro-ae3.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/intro-ae3.pdf")
```

In `rtf_*()` functions such as `rtf_body()`, `rtf_footnote()`,
the `text_justification` argument is used to align text.
Default is `"c"` for center justification.
To vary text justification by column, use character vector with length of vector equal to
number of columns displayed (e.g., `c("c", "l", "r")`).

All possible inputs can be found in the table below.

```{r}
r2rtf:::justification()
```

Below is an example to make the first column left-aligned and center-aligned for the rest.

```{r}
head(tbl) %>%
  rtf_body(text_justification = c("l", "c", "c", "c")) %>%
  rtf_encode() %>%
  write_rtf("tlf/intro-ae5.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/intro-ae5.pdf")
```

In `rtf_*()` functions such as `rtf_body()`, `rtf_footnote()`, etc.,
`border_left`, `border_right`, `border_top`, and `border_bottom` control cell borders.

If we want to remove the top border of `"Adverse Events"` in the header,
we can change the default value `"single"` to `""` in the `border_top` argument, as shown below.

`r2rtf` support 26 different border types. The details can be found on
the [r2rtf package website](https://merck.github.io/r2rtf/articles/rtf-row.html#border-type).

In this example, we also demonstrate the possibility of adding multiple column headers.

```{r}
head(tbl) %>%
  rtf_colheader(
    colheader = " | Treatment",
    col_rel_width = c(3, 6)
  ) %>%
  rtf_colheader(
    colheader = "Adverse Events | Placebo | Xanomeline High Dose | Xanomeline Low Dose",
    border_top = c("", "single", "single", "single"),
    col_rel_width = c(3, 2, 2, 2)
  ) %>%
  rtf_body(col_rel_width = c(3, 2, 2, 2)) %>%
  rtf_encode() %>%
  write_rtf("tlf/intro-ae7.rtf")
```

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("tlf/intro-ae7.pdf")
```

In the `r2rtf` R package [get started](https://merck.github.io/r2rtf/articles/r2rtf.html) page,
there are more examples to illustrate how to customize

- title, subtitle
- footnote, data source
- special character
- etc.

Those features will be introduced when we first use them in the rest of the chapters.