### 2022-03-18 images/pack.Rmd
Capture `pack()` output using asciicast

---
title: "Generate an asciicast SVG for pkglite::pack() output"
output: rmarkdown::html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  fig.path = "images/"
)
```

```{r echo=FALSE, results="hide"}
asciicast::init_knitr_engine(echo_input = FALSE)
```

```{r, include=FALSE}
options(asciicast_knitr_svg = TRUE)
```

```{asciicast, pack}
#' At: end
library("pkglite")

"esubdemo/" %>%
  collate(file_ectd(), file_auto("inst")) %>%
  pack(output = "r0pkgs.txt")
```