### 2022-03-23 index.Rmd

---
title: "R for Clinical Study Reports and Submission"
author: ["Yilong Zhang", "Nan Xiao", "Keaven Anderson"]
knit: "bookdown::render_book"
description: >
  This book will teach you how to prepare tables, listings, and figures for
  clinical study report and submit to regulatory agencies,
  the essential part of clinical trial development.

site: bookdown::bookdown_site

documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
---

```{r, include = FALSE}
library(bslib)
library(xml2)
library(downlit)
library(haven)
library(dplyr)
library(tidyr)
library(r2rtf)
library(knitr)
library(pkglite)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = "#>", echo = TRUE, message = FALSE)
options(dplyr.summarise.inform = FALSE)
```

```{r, eval = FALSE, include=FALSE}
r2rtf:::rtf_convert_format(
  list.files("tlf", pattern = ".rtf", full.names = TRUE),
  output_dir = "tlf"
)
```

```{r, eval = FALSE, include=FALSE}
rsconnect::writeManifest(appPrimaryDoc = "index.Rmd", contentCategory = "site")
```

```{r, eval = FALSE, include=FALSE}
bookdown::render_book("index.Rmd", "bookdown::pdf_book")
```

# Welcome! {-}

<img src="images/cover.jpg" width="100%" align="right" alt="" class="cover" />

Welcome to R for clinical study reports and submission.
Clinical study reports (CSR) are essential parts of clinical trial development.
A CSR is an "integrated" full scientific report of an individual clinical trial.

The [ICH E3: structure and content of clinical study reports](https://database.ich.org/sites/default/files/E3_Guideline.pdf)
provides guidance to assist sponsors in the development of a CSR.
In this book, you will learn how to use R to prepare a CSR and
how to submit it to regulatory agencies.

**This is a work-in-progress draft.**

## Events {-}

```{css, echo=FALSE}
table.table td:nth-child(1) { min-width: 27ch; }
table.table td:nth-child(2) { min-width: 12ch; }
table.table td:nth-child(3) { min-width: 10ch; }
table.table tbody { border-bottom: 1px solid #B6BDC6; }
```

```{r, echo=FALSE}
knitr::kable(read.table("events.tsv", stringsAsFactors = FALSE, header = TRUE, sep = "\t"))
```

## License {-}

This book is licensed to you under [Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License](https://creativecommons.org/licenses/by-nc-nd/4.0/).