### 2023-05-23 index.qmd

`r if (knitr::is_latex_output()) '<!--'`

# Welcome {.unnumbered}

Welcome to R for Clinical Study Reports and Submission.
Clinical study reports (CSR) are crucial components in clinical trial development.
A CSR is an "integrated" full scientific report of an individual clinical trials.

The [ICH E3: Structure and Content of Clinical Study Reports](https://database.ich.org/sites/default/files/E3_Guideline.pdf)
provides robust guidance to sponsors for developing a CSR.
This book demystifies the complexity by guiding you on how to use R for
streamlining the preparation of CSRs. Moreover, it walks you through the
process of preparing the submission to regulatory agencies. Regardless of your
expertise level in R programming, this book serves as an essential resource
in your statistical programming toolkit, enhancing efficiency and accuracy
in clinical trials reporting.

**This is a work-in-progress draft.**

## Events {.unnumbered}

```{r, echo=FALSE}
knitr::kable(read.table("events.tsv", stringsAsFactors = FALSE, header = TRUE, sep = "\t"))
```

## License {.unnumbered}

This book is licensed to you under the
[Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License](https://creativecommons.org/licenses/by-nc-nd/4.0/).

`r if (knitr::is_latex_output()) '-->'`