### 2023-06-04 project-folder.qmd
RStudio -> Posit except RStudio IDE

# Project folder {#sec-folder}

A clearly defined clinical project folder structure can have many benefits
to clinical development teams in an organization.
Specifically, a well-defined project structure can achieve:

- Consistency: everyone works on the same folder structure.
- Reproducibility: analysis can be executed and reproduced by different
  team members months/years later.
- Automation: automatically check the integration of a project.
- Compliance: reduce compliance issues.

We will use the [`esubdemo`](https://github.com/elong0527/esubdemo) R package to
illustrate the project folder structure for the A&R project. You can
[clone](https://happygitwithr.com/rstudio-git-github.html#clone-the-new-github-repository-to-your-computer-via-rstudio)
the project using RStudio IDE with

```bash
git clone https://github.com/elong0527/esubdemo.git
```

For R users, you already benefit from a well-defined and consistent folder structure.
That is the [R package folder structure](https://github.com/rstudio/cheatsheets/blob/main/package-development.pdf).
Every R package developer is required to follow the same convention to organize
their R functions before the R package can be disseminated through the Comprehensive R Archive Network (CRAN).
As a user, you can easily install and use those R packages after downloading from CRAN.
There are many good resources to guide developers on R package development,
such as, the [R Packages book](https://r-pkgs.org/) by Hadley Wickham.

We recommend using the R package folder structure to organize analysis scripts for clinical trial development.
Using the R package folder structure to streamline data analysis work has also been proposed before
(see @marwick2018packaging, @wuanalysis).

## Consistency {#sec-consistency}

For consistency, a well-defined folder structure with potential templates
ensures project teams organize the A&R work consistently across multiple projects.
Consistent folder structure also reduces communication costs
between study team members and enhances the transparency of projects.

In this book, we refer to an R package as a **project-specific R package**
if the purpose of an R package is to organize analysis scripts for a clinical project.

We refer to an R package as a **standard R package** if the purpose of an R package is
to share commonly used R functions to be hosted in a code repository such as CRAN.

Below is minimal sufficient folders and files for a project-specific R package
based on the R package folder structure.

- `*.Rproj`: RStudio project file used to open RStudio project.
- `DESCRIPTION`: Metadata for a package including authors, license, dependencies, etc.
- `R/`: Project-specific R functions.
- `vignettes/`: Analysis scripts using R Markdown.
- `man/`: Manual of project-specific R functions.

A general discussion of the R package folder structure can be found
in Chapter 4 of the R Packages book (@wickham2015r).

We demonstrate the idea using the [`esubdemo`](https://github.com/elong0527/esubdemo) project.

In the `esubdemo` project, we saved all TLF generation scripts in previous chapters into the `vignettes/` folder.

::: {.callout-note}
Under the `vignettes/` folder, there are two folders: `adam/` and `tlf/`.
The `adam/` folder contains ADaM datasets.
The `tlf/` folder contains output TLFs in RTF format.
We put `adam/` and `tlf/` folders within the `vignettes/` folder only for illustration purposes.
In an actual A&R report, you may have a different location to save your input and output.
:::

```
vignettes
├── data-adam
├── tlf
├── tlf-01-disposition.Rmd
├── tlf-02-population.Rmd
├── tlf-03-baseline.Rmd
├── tlf-04-efficacy.Rmd
├── tlf-05-ae-summary.Rmd
└── tlf-06-ae-spec.Rmd
```

While creating those analysis scripts, we also defined a few helper functions (e.g., `fmt_num` and `count_by`).
Those functions are saved in the `R/` folder.

```
R/
├── count_by.R
└── fmt.R
```

For a clinical trial project, it is also important to provide proper documentation for those help functions.
We use roxygen2 package to document functions.
For example, the header below defines each variable in `fmt_est`.
More details can be found in [Chapter 16 of the R Packages book](https://r-pkgs.org/man.html).

```{r, eval = FALSE}
#' Format point estimator
#'
#' @param .mean mean of an estimator.
#' @param .sd sd of an estimator.
#' @param digits number of digits for `.mean` and `.sd`.
#'
#' @export
fmt_est <- function(.mean, .sd, digits = c(1, 2)) {
  .mean <- fmt_num(.mean, digits[1], width = digits[1] + 4)
  .sd <- fmt_num(.sd, digits[2], width = digits[2] + 3)
  paste0(.mean, " (", .sd, ")")
}
```

The roxygen2 documentation will be converted into
standard R documentation format, and saved as `.Rd` files in the `man/` folder.
This step is automatically handled by `devtools::document()`.

```
man
├── count_by.Rd
├── fmt_ci.Rd
├── fmt_est.Rd
├── fmt_num.Rd
└── fmt_pval.Rd
```

The `man/` folder is used to save documentation automatically generated by roxygen2.
A typical workflow is to add roxygen2 documentation before each function in the `R/` folder.
Then `devtools::document()` is used to generate all the documentation files in the `man/` folder.
More details can be found in [Chapter 16 of the R Packages book](https://r-pkgs.org/man.html).

## Reproducibility {#sec-reproduce}

Reproducibility of analysis is one of the most important aspects of regulatory deliverables.
To ensure a successful reproduction, we need a controlled R environment,
including the control of the R version and the R package versions.
By using the R package folder structure and proper tools (e.g., renv, packrat),
we illustrate how to achieve reproducibility for R and R package versions.

::: {.callout-tip}
This is the same level of reproducibility in most SAS environments:
<https://support.sas.com/en/technical-support/services-policies.html#altos>
:::

### R version

First, we introduce the control of the R version.
In the esubdemo project, a reproducible environment is created when you open
the `esubdemo.Rproj` from RStudio IDE.
When we open the `esubdemo` project, RStudio IDE will execute the R code in `.Rprofile` automatically.
So we can use `.Rprofile` to set up a reproducible environment.
More details can be found in <https://rstats.wtf/r-startup.html>.
After we open the `esubdemo` project, the code in `.Rprofile` will
automatically check the current R version is the same as we defined in `.Rprofile`.

```{r, eval = FALSE}
# Set project R version
R_version <- "4.1.1"
```

If there is an R version mismatch, an error message is displayed as below.

```
Error: The current R version is not the same as the current project in R4.1.1
```

::: {.callout-caution}
`.Rprofile` is only for project-specific R packages.
A standard R package should not use `.Rprofile`.
:::

### R package version

Next, we introduce the control of the R package version, which is controlled in two layers.
Firstly, we define a snapshot date in `.Rprofile`.
The snapshot date allows us to freeze the source code repository.

```{r, eval = FALSE}
# set up snapshot date
snapshot <- "2021-08-06"

# set up repository based on the snapshot date
repos <- paste0("https://mran.microsoft.com/snapshot/", snapshot, "/")

# define repo URL for project-specific package installation
options(repos = repos)
```

We can also define the package repository to be a specific snapshot date.
For example, we used Microsoft R Application Network (MRAN) to
define the snapshot date to be `2021-08-06`.
The snapshot date freezes the R package repository.

In other words, all R packages installed in this RStudio project
are based on the frozen R version at the snapshot date.
Here it is `2021-08-06` by using the MRAN server.

Below information will be displayed after a new R session is opened.

```
Current project R package repository:
    https://mran.microsoft.com/snapshot/2021-08-06/
```

::: {.callout-note}
Posit Package Manager provides a solution to host both publicly available
and internally developed R packages. However, the
[Posit Public Package Manager](https://packagemanager.posit.co/)
does not provide a daily snapshot as MRAN does.
:::

Secondly, we use renv to lock R package versions and save them in the `renv.lock` file.
renv provides a robust and stable approach to managing R package versions
for project-specific R packages.
An introduction of renv can be found on its
[website](https://rstudio.github.io/renv/articles/renv.html).

```{r, eval = FALSE}
source("renv/activate.R")
```

The R code above in the `.Rprofile` initiates the renv running environment.
As a user, you can use `renv::init()`, `renv::snapshot()`, and `renv::restore()`
to initialize, save and restore R packages used for the current analysis project.

In the analysis project, the renv package will

- create a `renv.lock` file to save the state of package versions.
- create a `renv/` folder to manage R packages for a project.

::: {.callout-caution}
The `renv.lock` file and `renv/` folder are only for project-specific R package.
A standard R package should not use renv.
:::

In summary, the R package version is controlled in two layers.

- Define a snapshot date in `inst/startup.R`.
- Using renv to lock R versions within a project.

If the project is initiated properly,
you should be able to see similar messages to inform how we control R package versions.

```
* Project '~/esubdemo' loaded. [renv 0.14.0]
```

Once R packages have been properly installed,
the system will use the R packages located in the search path
defined based on the order of `.libPaths()`.
The startup message also provided the R package search path.

```
Below R package path are searching in order to find installed R packages in this R session:
    "/home/zhanyilo/github-repo/esubdemo/renv/library/R-4.1/x86_64-pc-linux-gnu"
    "/rtmp/RtmpT3ljoY/renv-system-library"
```

::: {.callout-tip}
A cloud-based R environment (e.g., [Posit Workbench](https://posit.co/products/enterprise/workbench/))
can enhance the reproducibility within an organization
by using the same operating system, R version, and R package versions for an A&R project.
More details can be found at <https://environments.rstudio.com/>.
:::

::: {.callout-note}
A container solution like Docker [@RJ-2020-007] could further enhance
the reproducibility across an organization at the operating system level
but beyond the scope of this book.
:::

In conclusion, to achieve reproducibility for a project-specific R package,
a clinical project team can work under a controlled R environment
in the same R version and R package versions defined by a repository snapshot date.

## Automation

By using the R package folder structure, you will benefit from many outstanding tools to
simplify and streamline your workflow.

We have learned a few functions in `devtools` to generate content automatically.
Here is a list of tools that can enhance the workflow.

- [devtools](https://devtools.r-lib.org/): make package development easier.
  + A good overview can be found in [Chapter 2 of the R Packages book](https://r-pkgs.org/Whole-game.html).
  + `devtools::load_all()`: load all functions in `R/` folder and running environment.
  + `devtools::document()`: automatically create documentation using roxygen2.
  + `devtools::check()`: automatically perform compliance check as an R package.
  + `devtools::build_site()`: automatically run analysis scripts in batch and create a pkgdown website.
- [usethis](https://usethis.r-lib.org/): automates repetitive tasks that arise during project setup and development.
- [testthat](https://testthat.r-lib.org/): streamline testing code.
  + A discussion of using the testthat for an A&R project can be found in (@madhutest).
- [pkgdown](https://pkgdown.r-lib.org/): generate static HTML documentation website for an R package
  + It also allows you to run all analysis code in batch.

You may further automatically execute routines by leveraging CI/CD workflow.
For example, the `esubdemo` project will rerun all required checks and build a pkgdown website
by using [Github Actions](https://usethis.r-lib.org/reference/github_actions.html).

As the consistent folder is defined, it also becomes easier to create specific tools
that fit the analysis and reporting purpose. Below are a few potential tools that can be helpful:

- Create project template using
  [RStudio project templates](https://rstudio.github.io/rstudio-extensions/rstudio_project_templates.html);
- Add additional compliance checks for analysis and reporting;
- Save log files for running in batch.

## Compliance

For a regulatory deliverable, it is important to maintain compliance.
With a consistent folder structure, we can define specific criteria for compliance.
Some compliance criteria can be implemented within the automatically checking steps.

For an R package, there are already criteria to ensure R package integrity.
More details can be found in [Chapter 20 of the R Packages book](https://r-pkgs.org/check.html).