### 2023-07-09 index.qmd
update index.qmd

`r if (knitr::is_latex_output()) '<!--'`

# Welcome {.unnumbered}

Welcome to R for Clinical Study Reports and Submission.
Clinical study reports (CSR) are crucial components in clinical trial development.
A CSR is an "integrated" full scientific report of an individual clinical trials.

The [ICH E3: Structure and Content of Clinical Study Reports](https://database.ich.org/sites/default/files/E3_Guideline.pdf)
offer comprehensive instructions to sponsors on the creation of a CSR.
This book is a clear and straightforward guide on using R to streamline the process of preparing CSRs. Additionally, it provides detailed guidance on the submission process to regulatory agencies. Whether you are a beginner or an experienced R programmer, this book is an indispensable asset in your clinical reporting arsenal.

**This is a work-in-progress draft.**

## Events {.unnumbered}

```{r, echo=FALSE}
knitr::kable(
  read.table("events.tsv", 
             stringsAsFactors = FALSE, 
             header = TRUE, 
             sep = "\t")
  )
```

## License {.unnumbered}

This book is licensed to you under the
[Creative Commons Attribution-NonCommercial-NoDerivatives 4.0 International License](https://creativecommons.org/licenses/by-nc-nd/4.0/).

`r if (knitr::is_latex_output()) '-->'`